In many, maybe most, situations, your dataset will have so many rows that it would not be realistic to plot every single observations.

This means that you will want to:

1.  Calculate some aggregate statisitcs first. Think of this "new" data object as just a new spreadsheet.
2.  Pass the smaller object (often a tibble or a data frame) into ggplot(). This the first layer.
3.  Choose the parts (rows and columns) which are most relevant.
4.  Experiment with...\
    ... arrangements within aes()\
    ... different geoms\
    ... sub-groups of the data (faceting may or may not be informative)\
    ... themes, labels, and so on.

# Toplines and crosstabs

```{r, echo=FALSE}
source("data_AJPS2021/0_ajps_recode.R")
```

```{r, echo=TRUE}
d1
```

```{r, echo=TRUE}
table(d2$climatechange)
```

```{r, echo=TRUE}
table(d2$climatechangeBIN)
```

```{r, echo=TRUE}
d2 %>% count(climatechangeBIN)
```

Are the missing observations the same for the original and the recoded variable? (If not, we would want to check whether earlier code did something unintended.)

```{r, echo=TRUE}
d2 %>% count(climatechangeBIN,climatechange)
```
