---
output: html_document
opts_chunk:
  message: FALSE
  warning: FALSE
editor_options: 
  chunk_output_type: console
---

# Economic data over time

```{r}
library(tidyverse)
D <- read_csv("data/macro/inflation_WDI.csv")
oecd <- read_csv("data/macro/oced_codes.csv")

#biggerText <- jzPack::biggerText

# Merge in OECD indicators
D <- left_join(D,oecd)
```

## Analysis of inflation patterns

In how many countries has inflation exceeded 10% in 2022:

```{r}
D %>% 
  filter(year == 2022) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  count(over10)
```

```{r}
D %>% 
  filter(year %in% c(2010:2022)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            over10 = sum(over10)) %>%
  ggplot(aes(x=year,y=over10)) +
  geom_col() +
  labs(y="", x="Year",
       title = "Number of countries with inflation 10%+",
       caption = "Data: WDI")
  
```

```{r}
D %>% 
  filter(year %in% c(2010:2022)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over10)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_col() +
  labs(y="", x="Year",
       title = "Proportion of countries where inflation exceeded 10%",
       caption = "Data: WDI")
```

There are some obvious problems with the x-axis here, so let's solve that, plus a few minor issues:

```{r}
barplot_V2 <-  D %>% 
  filter(year %in% c(2010:2022)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over10)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_col() +
  labs(y="", x="Year",
       title = "Proportion of countries where inflation exceeded 10%",
       caption = "Data: WDI") +
  scale_x_continuous(breaks = c(2010:2022)) +
  scale_y_continuous(labels = scales::percent) +
  theme(text = element_text(size=15)) +
  theme_bw() 

barplot_V2
```

We improved the labels of both axes, and increased the font size.

The new version is reasonably good, but there are still unnecessary gridlines, right?

You could either use `theme_classic()` instead of `theme_bw()`.

```{r}
barplot_V2 +
  theme_classic() 
```

Or we can keep `theme_bw()` and make some adjustments:

```{r}
barplot_V2 +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank())
```

```{r}
D %>% 
  filter(!is.na(inflation)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  ggplot(aes(x=year,
             y=over10)) +
  geom_col() +
  theme_bw() + 
  theme(text = element_text(size=14)) +
  labs(y="", x="Year",
       subtitle = "Number of countries where inflation exceeded 10%",
       caption = "Data: WDI.")
```

------------------------------------------------------------------------

### Zoom in on OECD countries and increase the time horizon

```{r}
D %>% filter(OECD==1) %>%
  filter(!is.na(inflation)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over10)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_col() +
  theme_bw() + 
  theme(text = element_text(size=15)) +
  scale_y_continuous(labels = scales::percent) +
  labs(y="", x="Year",
       title = "Proportion of countries where inflation exceeded 10%\n(OECD countries only)")
```

```{r}
D %>% 
  mutate(over5 = ifelse(inflation > 5, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over5)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_line() + geom_point() +
  theme_bw() + 
  theme(text = element_text(size=15)) +
  scale_y_continuous(labels = scales::percent) +
  labs(y="", x="Year",
       subtitle = "Global proportion of countries where inflation exceeded 5%",
       caption = "Data: WDI.")
```

```{r}
D %>% filter(OECD==1) %>%
  filter(!is.na(inflation)) %>%
  group_by(year) %>%
  summarise(M = median(inflation)) %>%
  ggplot(aes(x=year,
             y=M)) +
  # geom_point() + geom_line() +
  geom_col() +
  theme_bw() + 
  theme(text = element_text(size=15)) +
  labs(y="", x="Year",
       title = "Median inflation in OECD countries")
```

## Analysis of stock returns

Get data on stocks prices:

```{r}
#| eval: false
#| echo: true
library(tidyquant)
stocks <- tq_get(c("NVDA", "AVGO","^GSPC"),
                 from = "2022-01-15",
                 to = "2024-06-05")
```

```{r}
#| eval: true
#| echo: false
library(tidyverse)
library(tidyquant)
stocks <- read_csv("data/stocks2022-25.csv")
```

```{r}
stocks %>%
 ggplot(aes(x = date, y = adjusted, color = symbol)) +
 geom_line() +
 theme_minimal() +
 labs(title = "NVIDIA vs Broadcom Stock Prices",
      y = "Adjusted Price",
      color = "Stock")

```

This is not informative (we can only see what's going on for one of the series).

How about this?

```{r}
stocks %>%
 ggplot(aes(x = date, y = adjusted, color = symbol)) +
 geom_line() +
 facet_grid( ~ symbol)
```

Or perhaps this?

```{r}
stocks %>%
 ggplot(aes(x = date, y = adjusted, color = symbol)) +
 geom_line() +
 facet_wrap( ~ symbol,scales="free_y")
```


Slightly better, but do we really care about price? Price movements - returns and losses - are probably more important to investors.

So, calculate returns

```{r}
returns <- stocks %>%
 group_by(symbol) %>%
 tq_transmute(select = adjusted,
              mutate_fun = periodReturn,
              period = "daily",
              col_rename = "ret") %>%
 ungroup() %>%
 pivot_wider(names_from = symbol,
             values_from = ret) %>%
 na.omit()
```

Calculate rolling correlations:

```{r}
returns$roll_cor <- zoo::rollapply(data = returns[, c("NVDA", "AVGO")],
                              width = 30,
                              FUN = function(x) cor(x[,1], x[,2]),
                              by.column = FALSE,
                              align = "right",
                              fill = NA)

returns$roll_corNVSP <- zoo::rollapply(data = returns[, c("NVDA", "^GSPC")],
                                   width = 60,
                                   FUN = function(x) cor(x[,1], x[,2]),
                                   by.column = FALSE,
                                   align = "right",
                                   fill = NA)
```

30-Day Rolling Correlation of Daily Returns: NVIDIA vs Broadcom:

```{r}
returns %>%
 filter(date >= "2022-04-01") %>%
ggplot(aes(x = date, y = roll_cor)) +
 geom_line(color = "blue") +
 geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
 theme_minimal() +
 labs(subtitle = "30-Day Rolling Correlation of Daily Returns: NVIDIA vs Broadcom",
      title = "Do semiconductor industry giants move together?",
      y = "Correlation Coefficient",
      x = "Date") +
 scale_y_continuous(limits = c(-.5, 1))
```

```{r}
returns %>%
 filter(date >= "2022-04-01") %>%
 ggplot(aes(x = date, y = roll_corNVSP)) +
 geom_line(color = "blue") +
 geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
 theme_minimal() +
 labs(subtitle = "60-Day Rolling Correlation of Daily Returns: NVIDIA vs SP500",
      title = "Do semiconductor industry giants move together?",
      y = "Correlation Coefficient",
      x = "Date") +
 scale_y_continuous(limits = c(-.5, 1))
```

Scatterplots of returns

```{r}
ggplot(returns, aes(x = NVDA, y = AVGO)) +
 geom_point() +
 geom_smooth(method = "lm") +
 labs(title = "Scatterplot of Daily Returns",
      x = "NVIDIA",
      y = "Broadcom") +
 geom_abline(intercept = 0, slope = 1, 
             color = "red3",linetype=2,linewidth=1) 
```

```{r}
ggplot(returns, aes(x = NVDA, y = AVGO)) +
 geom_point() +
 labs(title = "Scatterplot of Daily Returns",
      x = "NVIDIA",
      y = "Broadcom")
```

```{r}
ggplot(returns, aes(x = `^GSPC`, y = AVGO)) +
 geom_point() +
 labs(title = "Scatterplot of Daily Returns",
      x = "S&P500",
      y = "Broadcom")
```

```{r}
ggplot(returns, aes(x = `^GSPC`, y = NVDA)) +
 geom_point() +
 labs(title = "Scatterplot of Daily Returns",
      x = "S&P500",
      y = "NVIDIA")
```

Let's also look at correlation during different market conditions

```{r}
returns %>%
 mutate(year = year(date)) %>%
 group_by(year) %>%
 summarize(correlation = cor(NVDA, AVGO)) %>%
 print()
```

```{r}
returns %>% mutate(ym = floor_date(date,"month")) %>%
 group_by(ym) %>%
 summarize(correlation = cor(NVDA, `^GSPC`)) %>%
 ggplot(aes(x = ym, y = correlation)) +
 geom_col()
 
```
