---
output: html_document
opts_chunk:
  message: FALSE
  warning: FALSE
editor_options: 
  chunk_output_type: console
---

# Economic data over time

```{r}
library(tidyverse)
D <- read_csv("data_macro/inflation_WDI.csv")
oecd <- read_csv("data_macro/oced_codes.csv")

#biggerText <- jzPack::biggerText

# Merge in OECD indicators
D <- left_join(D,oecd)
```

In how many countries has inflation exceeded 10% in 2022:

```{r}
D %>% 
  filter(year == 2022) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  count(over10)
```

```{r}
D %>% 
  filter(year %in% c(2010:2022)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            over10 = sum(over10)) %>%
  ggplot(aes(x=year,y=over10)) +
  geom_col() +
  labs(y="", x="Year",
       title = "Number of countries with inflation 10%+",
       caption = "Data: WDI")
  
```

```{r}
D %>% 
  filter(year %in% c(2010:2022)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over10)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_col() +
  labs(y="", x="Year",
       title = "Proportion of countries where inflation exceeded 10%",
       caption = "Data: WDI")
```

Ther are some obivous problems with the x-axis here, so let's solve that, plus a few minor issues:

```{r}
barplot_V2 <-  D %>% 
  filter(year %in% c(2010:2022)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over10)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_col() +
  labs(y="", x="Year",
       title = "Proportion of countries where inflation exceeded 10%",
       caption = "Data: WDI") +
  scale_x_continuous(breaks = c(2010:2022)) +
  scale_y_continuous(labels = scales::percent) +
  theme(text = element_text(size=15)) +
  theme_bw() 

barplot_V2
```

We improves the labels of both axies, and increased the font size.

The new version is reasonably good, but there are still unnecessary gridlines, right?

You could either use `theme_classic()` instead of `theme_bw()`.

```{r}
barplot_V2 +
  theme_classic() 
```

Or we can keep `theme_bw()` and make some adjustments:

```{r}
barplot_V2 +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank())
```


```{r}
D %>% 
  filter(!is.na(inflation)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  ggplot(aes(x=year,
             y=over10)) +
  geom_col() +
  theme_bw() + 
  theme(text = element_text(size=14)) +
  labs(y="", x="Year",
       subtitle = "Number of countries where inflation exceeded 10%",
       caption = "Data: WDI.")
```

------------------------------------------------------------------------

# Zomm in on OECD countries and increase the time horizon

```{r}
D %>% filter(OECD==1) %>%
  filter(!is.na(inflation)) %>%
  mutate(over10 = ifelse(inflation > 10, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over10)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_col() +
  theme_bw() + 
  theme(text = element_text(size=15)) +
  scale_y_continuous(labels = scales::percent) +
  labs(y="", x="Year",
       title = "Proportion of countries where inflation exceeded 10%\n(OECD countries only)")
```



```{r}
D %>% 
  mutate(over5 = ifelse(inflation > 5, 1, 0)) %>%
  group_by(year) %>%
  summarise(n = n(),
            prop = sum(over5)/n) %>% 
  ggplot(aes(x=year,
             y=prop)) +
  geom_line() + geom_point() +
  theme_bw() + 
  theme(text = element_text(size=15)) +
  scale_y_continuous(labels = scales::percent) +
  labs(y="", x="Year",
       subtitle = "Global proportion of countries where inflation exceeded 5%",
       caption = "Data: WDI.")
```



```{r}
D %>% filter(OECD==1) %>%
  filter(!is.na(inflation)) %>%
  group_by(year) %>%
  summarise(M = median(inflation)) %>%
  ggplot(aes(x=year,
             y=M)) +
  # geom_point() + geom_line() +
  geom_col() +
  theme_bw() + 
  theme(text = element_text(size=15)) +
  labs(y="", x="Year",
       title = "Median inflation in OECD countries")
```
