# Standard charts

We'll make bar charts, scatterplots, stacked overlapping distributions (joyplots), and some other charts, using data on which the paper *American Politics in Two Dimensions* is based:

```{r}
library(tidyverse)
library(haven)
library(labelled)

theme_set(theme_minimal())
theme_update(text = element_text(size=13),
             #text = element_text(family="Source Sans Pro")
)
             
# READ IN RECODED DATA
source("data_AJPS2021/0_ajps_recode.R")
```

As always, take a quick look at the structure of the datasets.

In this case, we have data from 3 surveys, which are stored in `d1`, `d2`, and `d3`.

```{r, echo=TRUE}
head(d2)
```

We'll start by looking at the distribution of a binary (and moralizing) view of politics; how do people respond to the prompt *"Politics is a battle between good and evil"*?

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100)
```

Now let's create a cross-tab, breaking down the responses by party ID:

A binary (and moralizing) view of politics

```{r}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100)
```

Let's plot the frequency of this Manichean perspective:

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent)) +
  geom_bar(position="dodge", stat="identity") + xlab("Percent") + ylab(NULL) +
  ggtitle("Opinion: Politics is a battle between good and evil")
```

Would it be useful to plot bars in different colors? Potetnially, but not like this...

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col()
```

Let's:

-   Change/improve the colors
-   Clean up the labels as appropriate

This scale might work - here darker colors indicate greater disagreement:

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col() +
  scale_fill_viridis_d() +
  labs(x="Proportion of respondents", y= "", fill = "",
       subtitle = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).")
```

## Faceting by party ID

Let's try to:

-   add `facet_wrap(~pid)`
-   which means we also have to use `group_by(pid)` before running `count()`

```{r}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col() + scale_fill_viridis_d() +
  labs(x="Proportion of respondents", y= "", fill = "", subtitle = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).") +
  facet_wrap(~pid) 
```

Also, we really need to show the party labels, not their numbers.

Here `as_factor(variable)` will work as long as `variable` is indeed labelled:

```{r}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col() + scale_fill_viridis_d() +
  labs(x="Proportion of respondents", y= "", fill = "", subtitle = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).") +
  facet_wrap(~as_factor(pid)) +
    theme(text=element_text(size=9))
```

Note that we often want to increase the size of the text elements (`theme(text=element_text(size=...))`) but in this case I'm actually making the text **smaller** so that facet labels fit on the page.

Are stronger partisans are more Manichean on average?

```{r goodevil-b}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(pid),x=percent,fill=as_factor(goodevil))) +
  geom_col() +
  scale_fill_viridis_d(alpha=.885) +
  labs(x="Proportion of respondents", y= "", fill = "",
      title = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).")
```

## Showing more variables together

Let's create several data objects: each of them will contain responses to the components of the conspiracy thinking scale:

```{r}
#| echo: true

pop2share <- d2 %>% 
              count(pop2) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(pop2)) %>%
              mutate(q = "People who have studied for a long time\nand have many diplomas do not really know\nwhat makes the world go round.")

officialshare <- d2 %>% 
              count(official) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(official)) %>%
              mutate(q = "Official government accounts of events\ncannot be trusted.")

con1share <- d2 %>% 
              count(con1) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(con1)) %>%
              mutate(q = "Even though we live in a democracy,\na few people will always run things anyway")

con4share <- d2 %>% 
              count(con4) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(con4)) %>%
              mutate(q = "Much of our lives are being controlled\nby plots hatched in secret places.")
```

Create one larger data objhect:

```{r}
shareShow <-
  bind_rows(
    pop2share,
    officialshare,
    con1share,
    con4share
  ) %>%
  filter(!is.na(categories))

head(shareShow)
```

Make a plot:

```{r}
#| fig-height: 7
#| fig-width: 12

shareShow %>%
  ggplot(aes(y=as_factor(q),x=percent,fill=as_factor(categories))) +
  geom_bar(position="stack", stat="identity", width = .5) +
  jcolors::scale_fill_jcolors(palette = "pal4") +
  theme_minimal() + theme(text = element_text(size=15)) +
  labs(y = "",x = "Percent", fill = "",title = "Uscinski et al. (AJPS, 2021) survey data\non anti-establishment sentiment")
```

Look at the components of `jcolors`.

```{r}
jcolors::display_all_jcolors()
```

## Putting anti-establishment thinking on the 2nd axis

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point() +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking") + theme_gray()
```

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point() +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking") +
  theme_bw()
```

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point() +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking") +
  theme_minimal()
```

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point(color = "purple", alpha=.55) +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking", title = "Conspiracy, populist, and Manichean orientations\nare orthogonal to the standard partisan divide", caption = "Data: Uscinski et al. (AJPS, 2021).") +
  theme_minimal()
```

## Conspiratorial orientation and PID

In principle, a box plot might make sense in this context...

```{r}
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) +
  theme_minimal() 
```

You could simultaneously display all respondets (jittered):

```{r}
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) +
  geom_jitter(color="purple",alpha=.4,width = .1) +
  theme_minimal() 
```

Make small edits:

```{r}
#| fig-align: center
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) + 
  geom_jitter(color="purple",alpha=.4,width = .1) + theme_minimal() +
  labs(y="Conspiratorial thinking\n(Average agreement with 4 questions)",x = "")
```

Perhaps even better:

```{r}
#| fig-align: center
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) + geom_jitter(color="purple",alpha=.4,width = .1) + theme_minimal() +
  labs(y="Conspiratorial thinking\n(Average agreement with 4 questions)",x = "",subtitle = "Conspiratorial thinking is uncorrelated with partisanship",
       caption = "Q1: Even though we live in a democracy, a few people will always run things anyway.
Q2: The people who really run the country, are not known to the voters.
Q3: Big events like wars, the recent recession, and the outcomes of elections are controlled\nby small groups of people who are working in secret against the rest of us.
Q4: Much of our lives are being controlled by plots hatched in secret places.")
```

## All 3 components of the anti-est. orienation

```{r}
#| fig-align: center
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=suspicion2)) +
  geom_boxplot(fill="purple", alpha=.2) +
  geom_jitter(alpha=.2,width = .11,color="purple4") +
  theme_minimal() +
  labs(y="Conspiratorial thinking + Populism + Manichean outlook",x = "")
```

## Narcissism as a correlate for conspiracy thinking?

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges()
```

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges(
    fill = "#00AFBB",
    quantile_lines = TRUE, quantiles = 2,alpha = .9,color = "white") +
  labs(y= "I tend to want others to admire me",x="Conspiracy thinking")
```

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges(
    fill = "#00AFBB",
    quantile_lines = TRUE, quantiles = 2,alpha = .9,color = "white") +
    xlim(0,1) + labs(y= "I tend to want others to admire me",x="Conspiracy thinking")
```

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges(
    fill = "#00AFBB",
    quantile_lines = TRUE, quantiles = 2,alpha = .9,color = "white") +
  xlim(0,1) + labs(y= "I tend to want others to admire me",
       x = "The horizontal dimension measures how strongly respondents exhibit 3 traits:
1. Conspiratorial thinking (e.g. \"Our lives are controlled by secret plots\")
2. Populist beliefs
3. Manichean political views",caption = "Data: Uscinski et al. (AJPS, 2021).")
```

```{r}
d1 %>%
  ggplot(aes(x=narcissism,y=suspicion2)) +
  geom_smooth() + labs(x = "Narcissism", y="Anti-establishment orientation") 
```

```{r}
d1 %>%
  ggplot(aes(x=narcissism,y=suspicion2)) +
  geom_smooth() + labs(x = "Narcissism", y="Anti-establishment orientation") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") 
```

## Mainstream news

```{r}
#| fig-height: 7
#| fig-width: 12
#| output: false
  
ggpubr::ggarrange(
  d1 %>%
  ggplot(aes(x=suspicion2,y=msm)) +
  geom_smooth() +
  labs(x = "Anti-establihment orientation", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") ,
  
d1 %>%
  ggplot(aes(x=narcissism,y=msm)) +
  geom_smooth() +
  labs(x = "Narcissism", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") 
)
```

```{r}
#| fig-height: 7
#| fig-width: 12
#| echo: false
  
ggpubr::ggarrange(
  d1 %>%
  ggplot(aes(x=suspicion2,y=msm)) +
  geom_smooth() +
  labs(x = "Anti-establihment orientation", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") ,
  
d1 %>%
  ggplot(aes(x=narcissism,y=msm)) +
  geom_smooth() +
  labs(x = "Narcissism", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") 
)
```

```{r}
collmod <- lm(msm ~ clintonft*suspicion2, data=d1)
ggeffects::ggeffect(collmod, terms=c("suspicion2","clintonft")) %>% plot() +
  labs(color="Rating of Hillary Clinton",y="Much of the mainstream news is deliberately slanted to mislead us",
       x="Anti-establishment orientation",title="")
```

"I often disagree with conventional views about the world"

```{r}
ggpubr::ggarrange(
d1 %>%
  ggplot(aes(x=suspicion2,y=conwis)) +
  geom_smooth() +
  labs(x = "Anti-establihment orientation", y="I often disagree with conventional views about the world") +
  ylim(c(1,5)),
d1 %>%
  ggplot(aes(x=narcissism,y=conwis)) +
  geom_smooth() +
  labs(x = "Narcissism", y="I often disagree with conventional views about the world") +
  ylim(c(1,5))
)
```

## Denial of climate change

```{r, echo=TRUE}
table(d2$climatechange)
```

```{r, echo=TRUE}
table(d2$climatechangeBIN)
```

```{r, echo=TRUE}
d2 %>% count(climatechangeBIN)
```

Are the missing observations the same for the original and the recoded variable? (If not, we would want to check whether earlier code did something unintended.)

```{r, echo=TRUE}
d2 %>% count(climatechangeBIN,climatechange)
```
