# Advanced ggplot

This chapter demonstrates how to build a polished, information-dense visualization using ggplot2's layered grammar of graphics. We'll create a **ridgeline plot** that reveals the relationship between psychological traits and conspiracy thinking.

## Setup and Data

```{r}
library(tidyverse)
source("data/ajps2021/0_ajps_recode.R")
```

The data comes from a study on conspiracy beliefs, containing survey responses where participants' psychological traits (incl. the "dark triad"---narcissism, Machiavellianism, and sociopathy---plus argumentativeness) can be estimated on the basis of questions that participants answered about themselves. 

## Selecting Variables for Analysis

We begin by selecting survey items that measure psychological traits. Each item asks respondents to rate their agreement with a statement on a scale:

```{r}
x <- c("argue1", "argue2", "argue3", "manip1", "manip2", "manip3", "manip4", 
       "attent1", "attent2", "attent3", "attent4",
       "insens1","insens2","insens3","insens4")

#d1$nondark <- ifelse(d1$insens3 <=2 & d1$manip4 <=2 & d1$insens1<=2, 1,0)
d1$nondark <- ifelse(d1$manipulate2 <=.2 & d1$sociopathy2 <= .2 & d1$attend2 <=.2, 1,0)

```

We also create a "nondark" indicator that flags respondents who score low on multiple dark triad dimensions, giving us a baseline comparison group.

## The Ridgeline Plot

Ridgeline plots (also called "joy plots") stack multiple density distributions vertically, making it easy to compare distributions across groups. Here, each ridge represents respondents who agreed with a particular psychological statement, and the x-axis shows their distribution of conspiracy thinking scores.

This visualization works well because it:

- **Shows distributions, not just averages**: We can see the full shape of conspiracy thinking for each group
- **Enables comparison**: Stacking the ridges vertically makes patterns immediately visible
- **Uses color strategically**: Grouping items by their parent concept (Narcissism, Machiavellianism, etc.) adds a second layer of information without clutter
- **Orders meaningfully**: Sorting by average conspiracy thinking puts the most revealing patterns at the extremes

```{r}
#| echo: false
#| eval: true
#| fig-width: 8
#| fig-height: 10
d1 %>% select(consp_Index2,
              nondark,
              all_of(x)) %>% 
  pivot_longer(cols = -consp_Index2) %>%
  mutate(name2 = case_when(
    name == "argue1" ~  "I like to argue online with other people",
    name == "argue2" ~  "I enjoy a good argument over a controversial issue",
    name == "argue3" ~  "I am willing to express my opinion online\neven if others strongly disagree with me",
    name == "manip1" ~  "I tend to manipulate others to get my way",
    name == "manip2" ~  "I have used deceit or lied to get my way",
    name == "manip3" ~  "I have used flattery to get my way",
    name == "manip4" ~  "I tend to exploit others towards my own end",
    name == "attent1" ~  "I tend to want others to admire me",
    name == "attent2" ~  "I tend to want others to pay attention to me",
    name == "attent3" ~  "I tend to seek prestige or status",
    name == "attent4" ~  "I tend to expect special favors from others",
    name == "insens1" ~ "I tend to lack remorse",
    name == "insens2" ~ "I tend to be unconcerned with the morality\nof my actions",
    name == "insens3" ~ "I tend to be callous or insensitive",
    name == "insens4" ~ "I tend to be cynical",
    name == "nondark" ~ "Low on narcissism, mach., and sociopathy"
  ),
  concept = case_when(
    name %in% c("argue1","argue2","argue3") ~ "Argumentativeness",
    name %in% c("manip1", "manip2", "manip3", "manip4") ~ "Machiavellianism",
    name %in% c("attent1", "attent2", "attent3", "attent4") ~ "Narcissism",
    name %in% c("insens1","insens2","insens3","insens4") ~ "Sociopathy",
    name %in% c("nondark") ~ "Non-dark"
  )) %>%
  filter(!is.na(value) & !is.na(consp_Index2)) %>%
  mutate(agree = ifelse(value>=5,1,0)) %>% 
  group_by(name) %>% mutate(avg_agree = mean(agree)) %>%
  group_by(name,agree) %>% mutate(avg_cons = mean(consp_Index2)) %>%
  ungroup() %>%
  filter(agree==1 | (name=="nondark" & value==1)) %>%
  ggplot(aes(x=consp_Index2,y=fct_reorder(name2,avg_cons),
             fill=concept)) +
  ggridges::geom_density_ridges() +
  theme_bw()  +
  theme(legend.position = "right") +
  labs(x="Conspiracy thinking",y="",color="",fill="Concept",
       title = "Distribution of conspiracy thinking\nbroken down by responses to psychological questions\nand color-coded by trait (dark triad + argumentativeness)") +
  #scale_fill_manual(values=Prism5) +
  see::scale_fill_social() +
  scale_x_continuous(limits = c(0,1))
```

## Key Technique: Reshaping Data with `pivot_longer()`

The most important step in making this visualization work is **reshaping the data from wide to long format**. In the original dataset, each psychological item is its own column. But ggplot2 works best when we have one observation per row, with a column indicating which variable we're looking at.

```{r}
d1 %>% select(consp_Index2,
              nondark,
              all_of(x)) %>% 
  pivot_longer(cols = -consp_Index2)
```

`pivot_longer()` transforms our data so that:

- Each respondent now has multiple rows (one per survey item they answered)
- The `name` column contains the variable name (e.g., "argue1", "manip2")
- The `value` column contains their response to that item

This "tidy" structure is what allows us to facet or group by the survey item---ggplot can now treat each item as a separate category on the y-axis.

## Building the Complete Visualization

The full pipeline combines several ggplot techniques to create a polished chart:

```{r}
#| echo: true
#| eval: false
d1 %>% select(consp_Index2,
              nondark,
              all_of(x)) %>% 
  pivot_longer(cols = -consp_Index2) %>%
  mutate(name2 = case_when(
    name == "argue1" ~  "I like to argue online with other people",
    name == "argue2" ~  "I enjoy a good argument over a controversial issue",
    name == "argue3" ~  "I am willing to express my opinion online\neven if others strongly disagree with me",
    name == "manip1" ~  "I tend to manipulate others to get my way",
    name == "manip2" ~  "I have used deceit or lied to get my way",
    name == "manip3" ~  "I have used flattery to get my way",
    name == "manip4" ~  "I tend to exploit others towards my own end",
    name == "attent1" ~  "I tend to want others to admire me",
    name == "attent2" ~  "I tend to want others to pay attention to me",
    name == "attent3" ~  "I tend to seek prestige or status",
    name == "attent4" ~  "I tend to expect special favors from others",
    name == "insens1" ~ "I tend to lack remorse",
    name == "insens2" ~ "I tend to be unconcerned with the morality\nof my actions",
    name == "insens3" ~ "I tend to be callous or insensitive",
    name == "insens4" ~ "I tend to be cynical",
    name == "nondark" ~ "Low on narcissism, mach., and sociopathy"
  ),
  concept = case_when(
    name %in% c("argue1","argue2","argue3") ~ "Argumentativeness",
    name %in% c("manip1", "manip2", "manip3", "manip4") ~ "Machiavellianism",
    name %in% c("attent1", "attent2", "attent3", "attent4") ~ "Narcissism",
    name %in% c("insens1","insens2","insens3","insens4") ~ "Sociopathy",
    name %in% c("nondark") ~ "Non-dark"
  )) %>%
  filter(!is.na(value) & !is.na(consp_Index2)) %>%
  mutate(agree = ifelse(value>=5,1,0)) %>% 
  group_by(name) %>% mutate(avg_agree = mean(agree)) %>%
  group_by(name,agree) %>% mutate(avg_cons = mean(consp_Index2)) %>%
  ungroup() %>%
  filter(agree==1 | (name=="nondark" & value==1)) %>%
  ggplot(aes(x=consp_Index2,y=fct_reorder(name2,avg_cons),
             fill=concept)) +
  ggridges::geom_density_ridges() +
  theme_bw()  +
  theme(legend.position = "right") +
  labs(x="Conspiracy thinking",y="",color="",fill="Concept",
       title = "Distribution of conspiracy thinking\nbroken down by responses to psychological questions\nand color-coded by trait (dark triad + argumentativeness)") +
  #scale_fill_manual(values=Prism5) +
  see::scale_fill_social() +
  scale_x_continuous(limits = c(0,1))
```

### What each step accomplishes

1. **Data selection and reshaping**: `select()` + `pivot_longer()` gives us one row per person-item combination

2. **Creating readable labels**: The `case_when()` inside `mutate()` replaces cryptic variable names (like "manip1") with the actual survey text ("I tend to manipulate others to get my way"). This makes the chart self-documenting.

3. **Adding concept groupings**: A second `case_when()` assigns each item to its parent psychological concept. This becomes our fill aesthetic for color-coding.

4. **Filtering to relevant subgroups**: We only show respondents who *agreed* with each statement (`agree == 1`), letting us ask: "Among people who say they manipulate others, what does their conspiracy thinking look like?"

5. **Smart ordering**: `fct_reorder(name2, avg_cons)` sorts the y-axis by average conspiracy thinking, so patterns emerge naturally---the traits most associated with conspiracy thinking appear at one end.

6. **Polished aesthetics**: `theme_bw()` provides a clean background, `see::scale_fill_social()` offers a friendly palette, and `labs()` adds informative titles.

The result is a chart that tells a story: psychological traits associated with the "dark triad" tend to cluster at higher levels of conspiracy thinking, while those low on all dark traits cluster are less conspiratorial. 

