# Heatmaps

This chapter demonstrates how to create **heatmaps** to display patterns across groups and time. We use Instagram posting data from Democratic presidential candidates during 2019-2020 to illustrate the technique.

## Setup and Data

```{r}
library(tidyverse)
library(lubridate)
```

We load Instagram posting data for US presidential candidates. This dataset contains individual posts with timestamps, engagement metrics, and account information.

```{r}
insta_pres <- read_csv("data/instagram_pres_candidates_2019_to_July162020.csv")
```

Let's examine the structure of the data:

```{r}
glimpse(insta_pres)
```

## Data Preparation

The key to creating effective heatmaps is aggregating data to the right level. For our time-based analysis, we need to extract month-year information from the posting dates.

```{r}
insta_pres <- insta_pres %>%
  mutate(
    date = as.Date(Created),
    monthyear = floor_date(date, "month")
  )
```

We create a `monthyear` variable by flooring each date to the first of its month. This gives us a consistent grouping variable for aggregation.

## Heatmaps with `geom_tile()`

Heatmaps display values in a matrix format where color intensity represents the magnitude of a variable. They're excellent for showing patterns across two categorical dimensions---in our case, candidates and time periods.

### Basic Heatmap: Selected Candidates

We start by filtering to five major candidates and counting posts per month:

```{r}
#| fig-width: 10
#| fig-height: 4
insta_pres %>% 
  filter(`User Name` %in% c("elizabethwarren", "joebiden", "berniesanders", 
                            "pete.buttigieg", "amyklobuchar")) %>%
  group_by(`User Name`, monthyear) %>%
  summarize(n = n(), .groups = "drop") %>% 
  ggplot(aes(monthyear, `User Name`, fill = n)) +
  geom_tile(color = "white", size = 0.1) + 
  theme_minimal() + 
  scale_fill_gradient(low = "white", high = "purple") +
  labs(x = "", y = "Instagram account", fill = "Number\nof posts",
       subtitle = "Instagram use of selected Democratic candidates")
```

Key elements of this visualization:

- **geom_tile()** creates rectangular tiles, one per candidate-month combination
- **color = "white"** adds thin borders between tiles for visual separation
- **scale_fill_gradient()** maps the count to a continuous color scale from white (low) to purple (high)

### Expanding to All Candidates

To show all candidates, we need to create a meaningful ordering. Here we order candidates by their total posting volume:

```{r}
# Create ordered list of accounts by total posts
account_order <- insta_pres %>%
  count(`User Name`) %>%
  arrange(desc(n)) %>%
  pull(`User Name`)
```

Now we can visualize all candidates:

```{r}
#| fig-width: 10
#| fig-height: 8
insta_pres %>% 
  group_by(`User Name`, monthyear) %>%
  summarize(n = n(), .groups = "drop") %>% 
  mutate(account = factor(`User Name`, levels = rev(account_order))) %>% 
  ggplot(aes(monthyear, account, fill = n)) +
  geom_tile(color = "white", size = 0.1) + 
  theme_minimal() + 
  scale_fill_gradient(low = "white", high = "navy") +
  labs(x = "", y = "Instagram account", fill = "Number\nof posts",
       subtitle = "Instagram use of Democrats who ran for president in 2019-2020")
```

### Categorical Color Scales

Sometimes a continuous color scale can be difficult to interpret, especially when the distribution is skewed. Converting counts to categories can make patterns more readable.

```{r}
library(RColorBrewer)
```

We use `cut()` to bin the count into meaningful categories:

```{r}
#| fig-width: 10
#| fig-height: 8
insta_pres %>% 
  filter(monthyear >= "2019-01-01") %>%
  group_by(`User Name`, monthyear) %>%
  summarize(n = n(), .groups = "drop") %>% 
  mutate(
    countfactor = cut(n, 
                      breaks = c(-1, 0, 5, 10, 20, 40, 60, 80, Inf),
                      labels = c("0", "1-5", "6-10", "11-20", "21-40", "41-60", "61-80", ">80"))
  ) %>%
  mutate(account = factor(`User Name`, levels = rev(account_order))) %>% 
  ggplot(aes(monthyear, account, fill = countfactor)) +
  geom_tile(color = "white", size = 0.1) + 
  theme_minimal() +
  scale_fill_manual(values = brewer.pal(8, "Purples"), na.value = "grey90") +
  labs(x = "", y = "Instagram account", fill = "Number\nof posts",
       subtitle = "Instagram use of Democrats who ran for president in 2019-2020")
```

Key differences in this approach:

- **cut()** creates bins with custom breakpoints---we chose bins that match how people naturally think about posting frequency
- **scale_fill_manual()** with `brewer.pal()` provides a sequential color palette designed for ordered categories
- **na.value = "grey90"** handles any missing month-candidate combinations

## Summary

This chapter covered heatmaps using `geom_tile()`:

- Best for showing values across two categorical dimensions (e.g., candidates and time periods)
- Use continuous color scales (`scale_fill_gradient()`) for fine-grained differences
- Use categorical color scales (via `cut()`) for easier interpretation of ranges
- Order categories meaningfully rather than alphabetically to reveal patterns
- Add white borders between tiles for visual separation

