---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Visualizing statistical models

A more accurate title, of course, woudl be "visualizing *outputs* from statistical models".

We'll be using this package to help us out:

```{r}
library(broom)
```

## Nationscape data

```{r}
# Read in Nationscape survey data:
library(tidyverse)
library(labelled)
a <- readRDS("data_nationscape2019/Nationscape_first10waves.rds")
```

### Association between sexism and AOC favorability over time

We estimate 10 separate model, aka "rolling regressions".

We run one model for each value of `week`, and create a tibble thanks to the `tidy()`function:

```{r}
a %>% filter(
         !is.na(aoc_Favorable),
         !is.na(gender_att3_by1SD)) %>%
  group_by(week) %>%
  do(broom::tidy(lm(aoc_Favorable ~
               gender_att3_by1SD +
               age + college_grad +
               White + Black + Hispanic, data = .))) %>%
    # Which coefficient we wish to pull:
    filter(term == "gender_att3_by1SD") %>%
    ggplot(aes(x=week,y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error)) +
  geom_pointrange(position = position_dodge(width = .45), size=.6) + labs(x="",y="Coefficient estimate",
  color="Subset of respondents") +
  ggtitle("Association between sexism and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) 
```

It seems that the correlation between sexism and (lower) favorability of AOC is quite stable.

```{r}
aoc_Favorable <- a %>% filter(
         !is.na(aoc_Favorable),
         !is.na(gender_att3_by1SD)) %>%
  group_by(pid3) %>%
  do(broom::tidy(lm(aoc_Favorable ~
               gender_att3_by1SD +
               age + college_grad +
               White + Black + Hispanic, data = .)))
```

### Demographics

We can now plot any set of coefficients, of course.

Let us try to see whether age correlates negatively with AOC's favorability, as we might perhaps suspect:

```{r}
aoc_Favorable %>% filter(pid3 <= 3 & term == "age") %>%
  ggplot(aes(x=as_factor(pid3),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,
             shape = factor(pid3))) +
  geom_pointrange() + labs(x="",y="Coefficient estimate",color="")
```

Let's improve the design a little bit:

```{r}
aoc_Favorable %>% filter(pid3 <= 3,term == "age") %>%
  ggplot(aes(x=as_factor(pid3),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,color=as_factor(pid3))) +
  geom_pointrange(position = position_dodge(width = .45), size=.6) + labs(x="Model estimate among respondents who identified as a...",y="Coefficient estimate",
  color="Subset of respondents") +
  ggtitle("Association between age and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) 
```


Need to fix the party colors:

```{r}
aoc_Favorable %>% filter(pid3 <= 3,term == "age") %>%
  ggplot(aes(x=as_factor(pid3),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,color=as_factor(pid3))) +
  geom_pointrange(position = position_dodge(width = .45), size=.6) + labs(x="Model estimate among respondents who identified as a...",y="Coefficient estimate",
  color="Subset of respondents") +
  ggtitle("Association between age and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) +
  scale_color_manual(values = c("blue3","darkred","grey40"))
```



## AJPS (2021) data

```{r}
library(tidyverse)
library(haven)
library(labelled)

# READ IN RECODED DATA
source("data_AJPS2021/0_ajps_recode.R")
```


We now turn again to the dataset posted by Uscinski et al. (2021).

```{r}
#| eval: true

# CT seems correlated with Trump FT among Dems
# Pop seems correlated with Trump FT among Reps 
# Manich. correlated with Trump FT among all

int1 <- lm(cexaggerate ~ pid*pop_Index, data=d3)
int2 <- lm(trumpft ~ pid*pop_Index, data=d3)
int3 <- lm(trumpft ~ pid*consp_Index, data=d3)
int4 <- lm(trumpft ~ pid*goodevil + consp_Index + pop_Index, data=d3)
```

```{r}
ggeffects::ggeffect(int1, terms=c("pid","pop_Index")) %>% plot() +
  labs(y="Climate change is exaggerated",
       x="PID (7=Strong Republican",color="Populism")
```




