---
title: "ggplot practice"
subtitle: "Context: American politics in 2 dimensions"
format: 
  revealjs:
    slide-number: c
execute:
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(haven)
library(labelled)
library(stargazer)


theme_set(theme_minimal())
theme_update(text = element_text(size=13),
             #text = element_text(family="Source Sans Pro")
)
             
# READ IN RECODED DATA
source("data_AJPS2021/0_ajps_recode.R")
```

## 

A binary (and moralizing) view of politics

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100)
```

## 

A binary (and moralizing) view of politics

```{r}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100)
```

##  {.smaller}

A binary (and moralizing) view of politics

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent)) +
  geom_bar(position="dodge", stat="identity") + xlab("Percent") + ylab(NULL) +
  ggtitle("Opinion: Politics is a battle between good and evil")
```

##  {.smaller}

A binary (and moralizing) view of politics

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col()
```

##  {.smaller}

A binary (and moralizing) view of politics

```{r}
d2 %>% 
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col() +
  scale_fill_viridis_d() +
  labs(x="Proportion of respondents", y= "", fill = "",
       subtitle = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).")
```

##  {.smaller}

A binary (and moralizing) view of politics by PID

```{r}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col() + scale_fill_viridis_d() +
  labs(x="Proportion of respondents", y= "", fill = "", subtitle = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).") +
  facet_wrap(~pid)
```

##  {.smaller}

A binary (and moralizing) view of politics by PID

```{r}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(goodevil),x=percent,fill=as_factor(goodevil))) +
  geom_col() + scale_fill_viridis_d() +
  labs(x="Proportion of respondents", y= "", fill = "", subtitle = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).") +
  facet_wrap(~as_factor(pid))
```

##  {.smaller}

Stronger partisans are more manichean on average

```{r goodevil-b}
d2 %>% group_by(pid) %>%
  count(goodevil) %>%
  mutate(percent = n/sum(n)*100) %>%
  ggplot(aes(y=as_factor(pid),x=percent,fill=as_factor(goodevil))) +
  geom_col() +
  scale_fill_viridis_d(alpha=.885) +
  labs(x="Proportion of respondents", y= "", fill = "",
      title = "Opinion: Politics is a battle between good and evil", caption= "Data: Uscinski et al. (AJPS, 2021).")
```

##  {.smaller}

```{r}
#| echo: true

pop2share <- d2 %>% 
              count(pop2) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(pop2)) %>%
              mutate(q = "People who have studied for a long time\nand have many diplomas do not really know\nwhat makes the world go round.")

officialshare <- d2 %>% 
              count(official) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(official)) %>%
              mutate(q = "Official government accounts of events\ncannot be trusted.")

con1share <- d2 %>% 
              count(con1) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(con1)) %>%
              mutate(q = "Even though we live in a democracy,\na few people will always run things anyway")

con4share <- d2 %>% 
              count(con4) %>%
              mutate(percent = n/sum(n)*100) %>%
              mutate(categories = as_factor(con4)) %>%
              mutate(q = "Much of our lives are being controlled\nby plots hatched in secret places.")

con4share
```

## 

```{r}
shareShow <-
  bind_rows(
    pop2share,
    officialshare,
    con1share,
    con4share
  ) %>%
  filter(!is.na(categories))

head(shareShow)
```

##  {.smaller}

```{r}
#| fig-height: 7
#| fig-width: 12

shareShow %>%
  ggplot(aes(y=as_factor(q),x=percent,fill=as_factor(categories))) +
  geom_bar(position="stack", stat="identity", width = .5) +
  jcolors::scale_fill_jcolors(palette = "pal4") +
  theme_minimal() + theme(text = element_text(size=15)) +
  labs(y = "",x = "Percent", fill = "",title = "Uscinski et al. (AJPS, 2021) survey data\non anti-establishment sentiment")
```

## 

```{r}
jcolors::display_all_jcolors()
```

##  {.smaller}

#### Putting anti-establishment thinking on the 2nd axis

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point() +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking") + theme_gray()
```

##  {.smaller}

#### Putting anti-establishment thinking on the 2nd axis

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point() +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking") +
  theme_bw()
```

##  {.smaller}

#### Putting anti-establishment thinking on the 2nd axis

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point() +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking") +
  theme_minimal()
```

##  {.smaller}

#### Putting anti-establishment thinking on the 2nd axis

```{r}
#| fig-height: 7
#| fig-width: 8
#| fig-align: center

d1 %>% ggplot(aes(x = leftright2, y = suspicion2)) +
  geom_point(color = "purple", alpha=.55) +
  labs(x = "Left vs. Right dimension\n(party ID, symbolic ideology, and party thermometer ratings)",
       y = "Anti-establishment thinking", title = "Conspiracy, populist, and Manichean orientations\nare orthogonal to the standard partisan divide", caption = "Data: Uscinski et al. (AJPS, 2021).") +
  theme_minimal()
```

## 

Conspiratorial orientation and PID

```{r}
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) +
  theme_minimal() 
```

## 

Conspiratorial orientation and PID

```{r}
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) +
  geom_jitter(color="purple",alpha=.4,width = .1) +
  theme_minimal() 
```

##  {.smaller}

Conspiratorial orientation and PID

```{r}
#| fig-align: center
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) + geom_jitter(color="purple",alpha=.4,width = .1) + theme_minimal() +
  labs(y="Conspiratorial thinking\n(Average agreement with 4 questions)",x = "")
```

##  {.smaller}

Conspiratorial orientation and PID

```{r}
#| fig-align: center
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=consp_Index)) +
  geom_boxplot(size=.4) + geom_jitter(color="purple",alpha=.4,width = .1) + theme_minimal() +
  labs(y="Conspiratorial thinking\n(Average agreement with 4 questions)",x = "",subtitle = "Conspiratorial thinking is uncorrelated with partisanship",
       caption = "Q1: Even though we live in a democracy, a few people will always run things anyway.
Q2: The people who really run the country, are not known to the voters.
Q3: Big events like wars, the recent recession, and the outcomes of elections are controlled\nby small groups of people who are working in secret against the rest of us.
Q4: Much of our lives are being controlled by plots hatched in secret places.")
```

##  {.smaller}

### All 3 components of the anti-est. orienation

```{r}
#| fig-align: center
d2 %>% 
  ggplot(aes(x=as_factor(pid), y=suspicion2)) +
  geom_boxplot(fill="purple", alpha=.2) +
  geom_jitter(alpha=.2,width = .11,color="purple4") +
  theme_minimal() +
  labs(y="Conspiratorial thinking + Populism + Manichean outlook",x = "")
```

## 

Narcissism as a correlate for conspiracy thinking?

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges()
```

##  {.smaller}

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges(
    fill = "#00AFBB",
    quantile_lines = TRUE, quantiles = 2,alpha = .9,color = "white") +
  labs(y= "I tend to want others to admire me",x="Conspiracy thinking")
```

##  {.smaller}

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges(
    fill = "#00AFBB",
    quantile_lines = TRUE, quantiles = 2,alpha = .9,color = "white") +
    xlim(0,1) + labs(y= "I tend to want others to admire me",x="Conspiracy thinking")
```

##  {.smaller}

```{r}
d1 %>% filter(!is.na(attent1)) %>%
  ggplot(aes(y = as_factor(attent1), x= suspicion2)) +
  geom_density_ridges(
    fill = "#00AFBB",
    quantile_lines = TRUE, quantiles = 2,alpha = .9,color = "white") +
  xlim(0,1) + labs(y= "I tend to want others to admire me",
       x = "The horizontal dimension measures how strongly respondents exhibit 3 traits:
1. Conspiratorial thinking (e.g. \"Our lives are controlled by secret plots\")
2. Populist beliefs
3. Manichean political views",caption = "Data: Uscinski et al. (AJPS, 2021).")
```

##  {.smaller}

```{r}
d1 %>%
  ggplot(aes(x=narcissism,y=suspicion2)) +
  geom_smooth() + labs(x = "Narcissism", y="Anti-establishment orientation") 
```

##  {.smaller}

```{r}
d1 %>%
  ggplot(aes(x=narcissism,y=suspicion2)) +
  geom_smooth() + labs(x = "Narcissism", y="Anti-establishment orientation") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") 
```

## 

Mainstream news

```{r}
#| fig-height: 7
#| fig-width: 12
#| output: false
  
ggpubr::ggarrange(
  d1 %>%
  ggplot(aes(x=suspicion2,y=msm)) +
  geom_smooth() +
  labs(x = "Anti-establihment orientation", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") ,
  
d1 %>%
  ggplot(aes(x=narcissism,y=msm)) +
  geom_smooth() +
  labs(x = "Narcissism", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") 
)
```

## 

Mainstream news

```{r}
#| fig-height: 7
#| fig-width: 12
#| echo: false
  
ggpubr::ggarrange(
  d1 %>%
  ggplot(aes(x=suspicion2,y=msm)) +
  geom_smooth() +
  labs(x = "Anti-establihment orientation", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") ,
  
d1 %>%
  ggplot(aes(x=narcissism,y=msm)) +
  geom_smooth() +
  labs(x = "Narcissism", y="Much of the mainstream news is deliberately slanted to mislead us") +
  ggside::geom_xsidehistogram() +
  ggside::ggside(x.pos = "bottom") 
)
```

## 

#### Mainstream news

```{r}
collmod <- lm(msm ~ clintonft*suspicion2, data=d1)
ggeffects::ggeffect(collmod, terms=c("suspicion2","clintonft")) %>% plot() +
  labs(color="Rating of Hillary Clinton",y="Much of the mainstream news is deliberately slanted to mislead us",
       x="Anti-establishment orientation",title="")
```

##  {.smaller}

I often disagree with conventional views about the world

```{r}
ggpubr::ggarrange(
d1 %>%
  ggplot(aes(x=suspicion2,y=conwis)) +
  geom_smooth() +
  labs(x = "Anti-establihment orientation", y="I often disagree with conventional views about the world") +
  ylim(c(1,5)),
d1 %>%
  ggplot(aes(x=narcissism,y=conwis)) +
  geom_smooth() +
  labs(x = "Narcissism", y="I often disagree with conventional views about the world") +
  ylim(c(1,5))
)
```

##  {.smaller}

```{r}
#| eval: false

# CT seems correlated with Trump FT among Dems
# Pop seems correlated with Trump FT among Reps 
# Manich. correlated with Trump FT among all

int1 <- lm(cexaggerate ~ pid*pop_Index, data=d3)
int2 <- lm(trumpft ~ pid*pop_Index, data=d3)
int3 <- lm(trumpft ~ pid*consp_Index, data=d3)
int4 <- lm(trumpft ~ pid*goodevil + consp_Index + pop_Index, data=d3)

ggeffects::ggeffect(int1, terms=c("pid","pop_Index")) %>% plot()
ggeffects::ggeffect(int1, terms=c("pop_Index","pid")) %>% plot()

ggeffects::ggeffect(int2, terms=c("pop_Index","pid")) %>% plot()
ggeffects::ggeffect(int3, terms=c("consp_Index","pid")) %>% plot()
ggeffects::ggeffect(int4, terms=c("goodevil","pid")) %>% plot()

library(rpart)

rpart(trumpft ~ pid3 + consp_Index + pop_Index + goodevil + edu + income + age, data = d3,
      control = rpart.control(cp = 0.003)) %>%
 rpart.plot::rpart.plot(., type=2, extra=1)
```

### Trump feeling thermometer

```{r}
library(rpart)
rpart::rpart(trumpft ~ pid3 + consp_Index + pop_Index + goodevil, data = d3,
      control = rpart.control(cp = 0.003)) %>%
 rpart.plot::rpart.plot(., type=2, extra=1)
```

Agnostic modeling:

-   Among Democrats, conspiracy thinking is prognostic.
-   Among Republicans, a Manichean view of politics is a useful predictor.

## 

### Birth certificate theory

#### Which variables are prognostic?

```{r}
lb0 <- glm(birtherBIN ~ suspicion2, data=d2constantSample, family = "binomial")
lb1 <- glm(birtherBIN ~ Republican, data=d2constantSample, family = "binomial")
lb2 <- glm(birtherBIN ~ trumpft2, data=d2constantSample, family = "binomial")
lb3 <- glm(birtherBIN ~ Republican + trumpft2, data=d2constantSample, family = "binomial")
lb4 <- glm(birtherBIN ~ Republican + trumpft2 + ideo2, data=d2constantSample, family = "binomial")
lb5 <- glm(birtherBIN ~ Republican + trumpft2 + ideo2 + suspicion2, data=d2constantSample, family = "binomial")

trb <- tribble(
  ~mod, ~PCP,
  "Anti-est. (only)",      getPCP(lb0),
  "PID (only)",      getPCP(lb1),
  "Trump FT (only)",      getPCP(lb2),
  "PID + Trump FT",      getPCP(lb3),
  "PID + Trump FT + Ideo."      ,getPCP(lb4),
  "PID + Trump FT + Ideo. + AEO",      getPCP(lb5)
)

trb %>%
  mutate(rowid = row_number()) %>%
  ggplot(aes(x=PCP,y=fct_reorder(mod,-rowid))) +
  geom_point() +
  #geom_col() +
  #scale_x_continuous(limits = c(70,88)) +
    labs(x="% Correctly predicted", y= "", title = "Opinion: Barack Obama faked his citizenship to become president")
```

```{r}
#| eval: false
l0 <- glm(climatechangeBIN ~ suspicion2, data=d2constantSample, family = "binomial")
l1 <- glm(climatechangeBIN ~ Republican, data=d2constantSample, family = "binomial")
l2 <- glm(climatechangeBIN ~ trumpft2, data=d2constantSample, family = "binomial")
l3 <- glm(climatechangeBIN ~ Republican + trumpft2, data=d2constantSample, family = "binomial")
l4 <- glm(climatechangeBIN ~ Republican + trumpft2 + ideo2, data=d2constantSample, family = "binomial")
l5 <- glm(climatechangeBIN ~ Republican + trumpft2 + ideo2 + suspicion2, data=d2constantSample, family = "binomial")

trb_climate <- tribble(
  ~mod, ~PCP,
  "Anti-est. (only)",      getPCP(l0),
  "PID (only)",      getPCP(l1),
  "Trump FT (only)",      getPCP(l2),
  "PID + Trump FT",      getPCP(l3),
  "PID + Trump FT + Ideo."      ,getPCP(l4),
  "PID + Trump FT + Ideo. + AEO",      getPCP(l5)
)

trb_climate %>%
  mutate(rowid = row_number()) %>%
  ggplot(aes(x=PCP,y=fct_reorder(mod,-rowid))) +
  geom_point() +
  #geom_col() +
  #scale_x_continuous(limits = c(70,88)) +
    labs(x="% Correctly predicted", y= "", title = "Opinion: Climate change is a hoax perpetrated by corrupt scientists and politicians")
```

```{r}
#| eval: false
list(
lm(collusion ~ consp_Index2, data = d2) ,
lm(collusion ~ goodevil2, data = d2) ,
lm(collusion ~ suspicion2 + 
     Democrat +
     Republican +
     college_grad + 
     female + 
     interest2, data = d2) 
) %>%
  modelsummary::modelsummary()
```
