---
title: "ggplot lesson"
subtitle: "Nationscape surveys"
format: 
  revealjs:
    slide-number: c
execute:
  echo: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(haven)
library(labelled)
library(pollster)
#library(stargazer)


theme_set(theme_minimal())
theme_update(text = element_text(size=13),
             #text = element_text(family="Source Sans Pro")
)
             
# READ IN RECODED DATA
a <- readRDS("data_nationscape2019/Nationscape_first10waves.rds")
```

## 

### List of variables

```{r}
dim(a)

names(a)
```

## We subset the data to the first 10 waves

```{r}
# Weekly number of respondents
a %>% group_by(week) %>% tally() %>% head()
```

. . .

```{r}
a %>% summarise(first = min(date), last = max(date))
```

## 

```{r}
a %>% group_by(week) %>% tally() %>%
  ggplot(aes(x=week,y=n)) + geom_bar(stat="identity") +
  labs(x="",y="Respondents",title="Weekly number of Nationscape respondents")
```

## 

```{r}
a %>% group_by(week) %>% tally() %>%
  ggplot(aes(x=week,y=n)) + geom_bar(stat="identity") +
  labs(x="",y="Respondents",title="Weekly number of Nationscape respondents") +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%W")
```

## 

```{r}
a %>% group_by(week) %>% tally() %>%
  ggplot(aes(x=week,y=n)) + geom_bar(stat="identity") +
  labs(x="",y="Respondents",title="Weekly number of Nationscape respondents") +
  scale_x_date(date_breaks = "1 months", minor_breaks = "1 weeks",date_labels = "%b %y")
```

## 

```{r}
a %>% group_by(week) %>% tally() %>%
  ggplot(aes(x=week,y=n)) + geom_line() + geom_point() +
  labs(x="",y="Respondents",title="Weekly number of Nationscape respondents") +
  scale_x_date(date_breaks = "1 months", minor_breaks = "1 days",date_labels = "%B %Y")
```

## 

```{r}
a %>% group_by(week) %>% tally() %>%
  ggplot(aes(x=week,y=n)) + geom_line() + geom_point() +
  labs(x="",y="Respondents",title="Weekly number of Nationscape respondents") +
  scale_x_date(date_breaks = "1 months", minor_breaks = "1 days",date_labels = "%B %Y") +
  ylim(c(0,8000))
```

## 

### Favorability of AOC

```{r}
table(a$aoc_Favorable)
```

## 

```{r}
a %>% summarise(AOC_fav = mean(aoc_Favorable))
```

. . .

```{r}
a %>% summarise(AOC_fav = mean(aoc_Favorable,na.rm=T),
                
                AOC_fav_w = weighted.mean(aoc_Favorable,
                                             w=weight,
                                             na.rm=T))
```

## Toplines vs. cross-tabs

Favorability of AOC

```{r}
pollster::topline(a,aoc_Favorable,weight = weight)
```

. . .

```{r}
pollster::crosstab(df = a, 
                        x = gender, 
                        y = aoc_Favorable,weight = weight) 
```

## 

```{r}
pollster::crosstab(df = a, x = gender, 
                        y = aoc_Favorable,weight = weight,format = "long") 
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") 
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1)
```

## Favorability by vote choice (intent)

```{r}
#| eval: false
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) 
# ggplot(aes(x= ..., y = ...))
```

## Favorability by vote choice (intent)

What if we had not specified `format = "wide"`?

```{r}
#| eval: true
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,
                   weight = weight)
```

## Favorability by vote choice (intent)

```{r}
#| eval: false
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(x= trump_biden, y = pct)) + geom_col()
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(x= trump_biden, y = pct)) + geom_col()
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(y= trump_biden, x = pct)) + geom_col()
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(y= trump_biden, x = pct)) + geom_col() + theme_classic()
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(y= fct_reorder(trump_biden,pct), 
             x = pct)) + geom_col() + theme_classic() + ggtitle("AOC favorability") 
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(y= fct_reorder(trump_biden,pct), 
             x = pct)) + geom_col() + theme_classic() +
  labs(x="Percent", y = "", title = "AOC favorability")
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(y= fct_reorder(trump_biden,pct), 
             x = pct)) + geom_col(width = .3,fill="darkorange") + theme_classic() +
  labs(x="Percent", y = "", title = "AOC favorability") 
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(y= fct_reorder(trump_biden,pct), 
             x = pct)) + geom_col(width = .3,fill="darkorange") + theme_classic() +
  labs(x="Percent", y = "", title = "AOC favorability") +
  geom_label(aes(label=pct))
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(y= fct_reorder(trump_biden,pct), 
             x = pct)) + geom_col(width = .3,fill="darkorange") + theme_classic() +
  labs(x="Percent", y = "", title = "AOC favorability") +
  geom_label(aes(label=round(pct,1)))
```

## Favorability by vote choice (intent)

```{r}
pollster::crosstab(df = a, x = trump_biden, 
                        y = aoc_Favorable,weight = weight,format = "long") %>%
  filter(aoc_Favorable==1, trump_biden!="Don't Know") %>%
  ggplot(aes(y= fct_reorder(trump_biden,pct), 
             x = pct)) + geom_col(width = .3,fill="darkorange") + theme_classic() +
  labs(x="Percent", y = "Supporters of...", title = "AOC favorability") +
  geom_label(aes(label=round(pct,1)))
```

# From 2-way to 3-way crosstabs

## 

Break down vote choice by education (later also by month)

```{r}
moe_crosstab(df= a %>% filter(registered==1,trump_biden != 999),
                  x = educ_category,
                  y = trump_biden,weight = weight,format="long",n=F,unwt_n=TRUE)
```

##  {.smaller}

```{r}
#| fig-alight: center
moe_crosstab(df= a %>% filter(registered==1,trump_biden != 999),
                  x = educ_category,
                  y = trump_biden,weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  ggplot(aes(x=educ_category,y=pct,fill=trump_biden)) + 
  geom_col() + scale_fill_brewer(palette = "Set1",direction = -1) +
  labs(x="",y="",fill="")
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a %>% filter(registered==1,trump_biden != 999),
                  x = educ_category,
                  y = trump_biden,z=monthyear,
                  weight = weight,n=F,unwt_n=TRUE) %>% head()
```

. . .

```{r}
moe_crosstab_3way(df= a %>% filter(registered==1,trump_biden != 999),
                  x = educ_category,
                  y = trump_biden,z=monthyear,
                  weight = weight,format="long",n=F,unwt_n=TRUE)
```

##  {.smaller}

```{r}
#| echo: false
moe_crosstab_3way(df= a %>% filter(registered==1,trump_biden != 999),
                  x = educ_category,
                  y = trump_biden,z=monthyear,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  ggplot(aes(x=monthyear,y=pct,color=trump_biden)) + 
  geom_line(lwd=.5) + geom_point(size=1.5) +
  geom_errorbar(aes(ymin = (pct - moe), ymax = (pct + moe),
                    color=trump_biden), width=.1) +
  labs(color="Preferred candidate",y="Two-party preference share",x="",
       title="2020 vote intent among white non-Hispanic registered voters",
       subtitle="Democracy Fund + UCLA Nationscape data") +
  theme_minimal() + facet_grid(~educ_category) +
  scale_color_brewer(palette = "Set1",direction = -1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=.5)) +
  scale_x_date(date_breaks = "1 month",date_labels = "%b 20%y")
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a %>% filter(registered==1,trump_biden != 999),
                  x = educ_category,
                  y = trump_biden,z=monthyear,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  ggplot(aes(x=monthyear,y=pct,color=trump_biden)) + 
  geom_line() + geom_point() +
  geom_errorbar(aes(ymin = (pct - moe), ymax = (pct + moe),
                    color=trump_biden)) +
  theme_minimal() + facet_grid(~educ_category)
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a %>% filter(registered==1,trump_biden != 999),
                  x = educ_category,
                  y = trump_biden,z=monthyear,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  ggplot(aes(x=monthyear,y=pct,color=trump_biden)) + 
  geom_line(lwd=.5) + geom_point(size=1.5) +
  geom_errorbar(aes(ymin = (pct - moe), ymax = (pct + moe), color=trump_biden), width=.1) +
  labs(color="Preferred candidate",y="Two-party preference share",x="",title="2020 vote intent among white non-Hispanic registered voters", subtitle="Democracy Fund + UCLA Nationscape data") +
  scale_color_brewer(palette = "Set1",direction = -1) + theme_minimal() + facet_grid(~educ_category) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=.5)) +
  scale_x_date(date_breaks = "1 month",date_labels = "%b 20%y")
```

## New target chart

```{r}
#| echo: false
#| fig-width: 6
#| fig-align: center

moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1,trump_biden!="Don't Know") %>%
  mutate(who = glue::glue('{trump_biden} voters')) %>%
  ggplot(aes(x=who,y=pct,color=age_group)) + 
  geom_point(size=2.5, position=position_dodge(width=.1)) +
  labs(color="Respondent age group",y="Percent with favorable views",x="",
       title="AOC favorability by vote choice and by age") +
  theme_minimal() + scale_color_brewer(palette = "Set2") +
  theme(legend.position = "top") + geom_hline(yintercept = 0,lty=2) +
  theme(panel.grid.major.x = element_blank(),panel.grid.minor.x = element_blank(), # remove the vertical grid lines
        panel.grid.major.y = element_blank()) +
  theme(text= element_text(size=13)) +
  guides(color = guide_legend(nrow = 2))
```

##  {.smaller}

### Favorability of AOC by age and vote intent

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE)
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1)
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(x=trump_biden,y=pct,color=age_group)) + 
  geom_point()
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1) %>%
  ggplot(aes(x=trump_biden,y=pct,color=age_group)) + 
   geom_point(size=2.5, position=position_dodge(width=.1))
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1,trump_biden!="Don't Know") %>%
  ggplot(aes(x=trump_biden,y=pct,color=age_group)) + 
  geom_point(size=2.5, position=position_dodge(width=.1))
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1,trump_biden!="Don't Know") %>%
  ggplot(aes(x=trump_biden,y=pct,color=age_group)) + 
  geom_point(size=2.5, position=position_dodge(width=.1)) +
  labs(color="Respondent age group",y="Percent with favorable views",x="",
       title="AOC favorability by vote choice and by age") +
  theme_minimal() + scale_color_brewer(palette = "Set2") 
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1,trump_biden!="Don't Know") %>%
  mutate(who = glue::glue('{trump_biden} voters')) %>%
  ggplot(aes(x=who,y=pct,color=age_group)) + geom_point(size=2.5, position=position_dodge(width=.1)) +
  labs(color="Respondent age group",y="Percent with favorable views",x="",title="AOC favorability by vote choice and by age") +
  theme_minimal() + scale_color_brewer(palette = "Set2") +
  guides(color = guide_legend(nrow = 2))
```

##  {.smaller}

```{r}
moe_crosstab_3way(df= a,x = age_group,y = aoc_Favorable,z=trump_biden,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  filter(aoc_Favorable==1,trump_biden!="Don't Know") %>%
  mutate(who = glue::glue('{trump_biden} voters')) %>%
  ggplot(aes(x=who,y=pct,color=age_group)) +  geom_point(size=2.5, position=position_dodge(width=.1)) +
  labs(color="Respondent age group",y="Percent with favorable views",x="", title="AOC favorability by vote choice and by age") +
  theme_minimal() + scale_color_brewer(palette = "Set2") +
  theme(legend.position = "top") + geom_hline(yintercept = 0,lty=2) +
  theme(panel.grid.major.x = element_blank(),panel.grid.minor.x = element_blank(), # remove the vertical grid lines
        panel.grid.major.y = element_blank()) +
  theme(text= element_text(size=13)) +
  guides(color = guide_legend(nrow = 2))
```

# Visualizing model coefficients

## 

### Favorability of AOC: a set of models

```{r}
library(broom)
aoc_Favorable <- a %>% filter(
         !is.na(aoc_Favorable),
         !is.na(gender_att3_by1SD)) %>%
  group_by(pid3) %>%
  do(tidy(lm(aoc_Favorable ~
               gender_att3_by1SD +
               age + college_grad +
               White + Black + Hispanic, data = .)))
```

##  {.smaller}

```{r}
aoc_Favorable
```

##  {.smaller}

```{r}
aoc_Favorable %>% 
  ungroup() %>%
  count(term)
```

##  {.smaller}

```{r}
aoc_Favorable %>%
  filter(pid3 <= 3,
         term == "age") %>%
  ggplot(aes(x=pid3,
             y=estimate,
             color=as_factor(pid3))) +
  geom_point()
```

##  {.smaller}

```{r}
aoc_Favorable %>%
  filter(pid3 <= 3,
         term == "age") %>%
  ggplot(aes(x=pid3,
             y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,
             color=as_factor(pid3))) +
  geom_pointrange()
```

## Use as_factor() {.smaller}

```{r}
aoc_Favorable %>% filter(pid3 <= 3,term == "age") %>%
  ggplot(aes(x=as_factor(pid3),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,
             color=as_factor(pid3))) +
  geom_pointrange() + labs(x="",y="Coefficient estimate",color="")
```

##  {.smaller}

```{r}
aoc_Favorable %>% filter(pid3 <= 3,term == "age") %>%
  ggplot(aes(x=as_factor(pid3),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,
             color=as_factor(pid3))) +
  geom_pointrange() + labs(x="",y="Coefficient estimate",color="") +
  geom_hline(yintercept=0) 
```

##  {.smaller}

```{r}
aoc_Favorable %>% filter(pid3 <= 3,term == "age") %>%
  ggplot(aes(x=as_factor(pid3),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,
             color=as_factor(pid3))) +
  geom_pointrange() + labs(x="",y="Coefficient estimate",color="") +
  geom_hline(yintercept=0, linetype=2) 
```

##  {.smaller}

```{r}
aoc_Favorable %>% filter(pid3 <= 3,term == "age") %>%
  ggplot(aes(x=as_factor(pid3),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error,color=as_factor(pid3))) +
  geom_pointrange(position = position_dodge(width = .45), size=.6) + labs(x="",y="Coefficient estimate",
  color="Subset of respondents") +
  ggtitle("Association between age and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) 
```

##  {.smaller}

Run regressions by ethnicity

```{r}
aoc_Fav_by_eth <- a %>% filter(
         !is.na(aoc_Favorable),
         !is.na(race_ethnicity)) %>%
  group_by(race_ethnicity) %>%
  do(tidy(lm(aoc_Favorable ~
               gender_att3_by1SD + factor(pid3) +
               age + college_grad, data = .)))
```

##  {.smaller}

Run regressions by ethnicity

```{r}
aoc_Fav_by_eth
```

##  {.smaller}

Display a subset of results

```{r}
aoc_Fav_by_eth %>% filter(race_ethnicity==2)
```

##  {.smaller}

```{r}
aoc_Fav_by_eth %>% filter(term == "age",race_ethnicity<=10) %>%
  ggplot(aes(x=as_factor(race_ethnicity),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error)) +
  geom_pointrange(size=1.5) + labs(x="",y="Coefficient estimate") +
  ggtitle("Association between age and AOC favorability")
```

##  {.smaller}

```{r}
aoc_Fav_by_eth %>% filter(term == "age",race_ethnicity<=10) %>%
  ggplot(aes(x=as_factor(race_ethnicity),y=estimate,
             ymax = estimate + 1.96*std.error,
             ymin = estimate - 1.96*std.error)) +
  geom_pointrange() + labs(x="",y="Coefficient estimate") +
  ggtitle("Association between age and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.5))
```

##  {.smaller}

```{r}
aoc_Fav_by_eth %>% filter(term == "college_grad",race_ethnicity<=10) %>%
  ggplot(aes(x=as_factor(race_ethnicity),y=estimate,
             ymax = estimate + 1.96*std.error,ymin = estimate - 1.96*std.error)) +
  geom_pointrange() + labs(x="",y="Coefficient estimate") +
  ggtitle("Association between education and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.5))
```

##  {.smaller}

```{r}
aoc_Fav_by_eth %>% filter(term == "factor(pid3)2",race_ethnicity<=10) %>%
  ggplot(aes(x=as_factor(race_ethnicity),y=estimate,
             ymax = estimate + 1.96*std.error,ymin = estimate - 1.96*std.error)) +
  geom_pointrange() + labs(x="",y="Coefficient estimate") +
  ggtitle("Association between Republican Party ID and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.5))
```

##  {.smaller}

```{r}
aoc_Fav_by_eth %>% filter(term == "gender_att3_by1SD",race_ethnicity<=10) %>%
  ggplot(aes(x=as_factor(race_ethnicity),y=estimate,
             ymax = estimate + 1.96*std.error,ymin = estimate - 1.96*std.error)) +
  geom_pointrange() + labs(x="",y="Coefficient estimate") +
  ggtitle("Association between sexist attitudes ID and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.5))
```

##  {.smaller}

```{r}
aoc_Fav_by_eth %>% filter(term == "gender_att3_by1SD",race_ethnicity<=10) %>%
  ggplot(aes(x=as_factor(race_ethnicity),y=estimate,
             ymax = estimate + 1.96*std.error,ymin = estimate - 1.96*std.error)) +
  geom_pointrange() + labs(x="",y="Coefficient estimate") +
  ggtitle("Association between sexist attitudes ID and AOC favorability") +
  geom_hline(yintercept=0, linetype=2) + 
  coord_flip()
```

# Heatmaps

##  {.smaller}

Immigration and vote choice

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 6
#| fig-align: center

pollster::crosstab_3way(df = a,
                        x = educ_category,
                        z = immig_quintile,
                        y = trump16,
                        weight = weight) %>%
  ggplot(aes(x = as_factor(immig_quintile),
             y = educ_category,
             fill = `1`, label = round(`1`,1))) +
  geom_tile(alpha=.77) +
  ggfittext::geom_fit_text(place = "topright",size=13) +
  jcolors::scale_fill_jcolors_contin(palette = "pal12") +
  labs(fill = "% Trump\nvote",
       x = "",
       y = "",
       title = "Trump vote in 2016 by preferences on immigration") +
  theme_minimal() +
  theme(text = element_text(size=15)) +
  theme(axis.text.x = element_text(angle = 40, vjust = 0.5, hjust=.5))
```

##  {.smaller}

The underlying data:

```{r}
pollster::crosstab_3way(df = a,
                        x = educ_category,
                        z = immig_quintile,
                        y = trump16,
                        weight = weight)
```

##  {.smaller}

```{r}
pollster::crosstab_3way(df = a,
                        x = educ_category,
                        z = immig_quintile,
                        y = trump16,
                        weight = weight) %>%
  ggplot(aes(x = as_factor(immig_quintile),y = educ_category,
             fill = `1`)) + geom_tile()
```

##  {.smaller}

```{r}
pollster::crosstab_3way(df = a,
                        x = educ_category,
                        z = immig_quintile,
                        y = trump16,
                        weight = weight) %>%
  ggplot(aes(x = as_factor(immig_quintile),y = educ_category,
             fill = `1`)) + geom_tile() +
  jcolors::scale_fill_jcolors_contin(palette = "pal12") +
  labs(fill = "% Trump\nvote",x = "",y = "",
       title = "Trump vote in 2016 by preferences on immigration")
```

##  {.smaller}

```{r}
pollster::crosstab_3way(df = a,
                        x = educ_category,
                        z = immig_quintile,
                        y = trump16,
                        weight = weight) %>%
  ggplot(aes(x = as_factor(immig_quintile),y = educ_category,
             fill = `1`, label = round(`1`,1))) +
  geom_tile(alpha=.77) +
  ggfittext::geom_fit_text(place = "topright",size=13) +
  jcolors::scale_fill_jcolors_contin(palette = "pal12") +
  labs(fill = "% Trump\nvote",x = "",y = "", title = "Trump vote in 2016 by preferences on immigration")
```

##  {.smaller}

```{r}
#| fig-align: center
pollster::crosstab_3way(df = a,
                        x = educ_category,
                        z = immig_quintile,
                        y = trump16,
                        weight = weight) %>%
  ggplot(aes(x = as_factor(immig_quintile),
             y = educ_category,
             fill = `1`, label = round(`1`,1))) +
  geom_tile(alpha=.77) +
  ggfittext::geom_fit_text(place = "topright",size=13) +
  jcolors::scale_fill_jcolors_contin(palette = "pal12") +
  labs(fill = "% Trump\nvote",x = "",y = "",title = "Trump vote in 2016 by preferences on immigration") +
  theme_minimal() + theme(text = element_text(size=15)) +
  theme(axis.text.x = element_text(angle = 40, vjust = 0.5, hjust=.5))
```

## 

```{r}
# a %>%
#   filter(extra_group_id_being_american <=4,
#          extra_group_identity_partyID <=4) %>%
#   count(extra_group_id_being_american,
#         extra_group_identity_partyID) %>%
#   mutate(prop = n / sum(n), cs = cumsum(prop)) %>%
#   ggplot(aes(x = as_factor(extra_group_id_being_american),
#              y = as_factor(extra_group_identity_partyID),
#              fill = prop* 100 )) +
#     geom_tile() +
#     scale_fill_gradient(low = "white", high = "darkblue") +
#     labs(fill = "% of respondents",
#          y = "Partisan Identity importance",
#          x = "American Identity importance",
#          caption = "Data: Nationscape")
```
